---
title: "assignment-4-task-2"
author: "nicquinn"
date: "2022-09-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(tidyr)
```

## R Markdown

First, the opportunity to read in the data set including occupational involvement between genders in 1870.

```{r}

occupation <- read_csv("occupation-1870.csv")
```

This data set can be gathered with the tidyr package tool, pivot_longer. The industry column was separated to reflect the different genders.

```{r}
longerData <- occupation %>%
  pivot_longer(c(2:11), names_to = "industry", values_to = "people_count") %>%
  separate(industry, into = c("industry", "gender"), "[.]") %>%
  rename(area_name = `Area name`) %>%
  rename(total_population = Total.Population)

print(head(longerData))

```

This data was then cleaned to make it easier to represent on a scatter plot.

```{r}
graphData <- longerData %>%
  select(industry, gender, people_count) %>%
  group_by(industry, gender) %>%
  summarise(industry_total = sum(people_count))

print(head(graphData))
```

Now, this data can be graphed utilizing the GGPLOT package from the tidyverse.

```{r}
graph1 <- ggplot(graphData, aes(industry, industry_total))

graph1 + geom_point(aes(colour = factor(gender))) +
  labs(x = "Occupation", y = "Amount of People Involved",
       title = "Occupational Gender Split in 1870") +
  theme(plot.title=element_text(hjust=0.5)) +
  scale_y_continuous(labels = function(x) format(x, scientific = FALSE), breaks = seq(0, 6000000, len = 6))
 
  
  
```

\ It is now graphically clear the variance between genders in each occupation. 